---
title: "Wk4_Assignment"
author: "Ryan Paw"
date: "10/22/2020"
output: html_document
---

```{r setup, include=FALSE, paged.print=FALSE}
knitr::opts_chunk$set(echo = TRUE)

myDataLocation <- "/Users/RyanPaw/My Files/R/ANA605/wk_4"

#SET WORKING DIRECTORY TO LOCATION OF DATA FILE
setwd(myDataLocation)

#IMPORT DATA AND PUT INTO DATAFRAME
dat <- read.csv(file = "admission_data.csv", header = TRUE)

# ========================
library(supernova)
library(mosaic)
library(ggplot2)
library(ggformula)
```

## Q1

1.	For each of the following variables contained in the below table,
    a.	Get the frequencies for each categorical variable and means/sd for each quantitative variable.

Variables
- Research Experience		
- Undergraduate GPA		
- Letter of Recommendation (LOR)		
- Statement of Purpose (SOP)	

b.	Get correlations (and p-values) between the outcome, chance of admit, and explanatory variable. Put those values in the table below, labeled r (p). 
c.	Perform a multiple regression analysis for the outcome variable, chance of admit, and fill in the rest of the table below.


```{r paged.print=FALSE}
str(dat)
#q1 a)
tally(~ Research, dat)
favstats(dat$CGPA)
favstats(dat$LOR)
favstats(dat$SOP)
tally(~SOP, dat)
```
```{r paged.print=FALSE}
#Q1 b
dat$ResearchF <- factor(dat$Research, levels=c(0,1), c("No","Yes"))

#Corr and P values
cor.test(Chance.of.Admit~Research,data=dat,conf.level=.95)
cor.test(Chance.of.Admit~CGPA,data=dat,conf.level=.95)
cor.test(Chance.of.Admit~LOR,data=dat,conf.level=.95)
cor.test(Chance.of.Admit~SOP,data=dat,conf.level=.95)
```
```{r paged.print=FALSE}
multi.model <- lm(Chance.of.Admit ~ CGPA + SOP + LOR + Research, data=dat)
multi.model
summary(multi.model)
supernova(multi.model)
confint(multi.model)
```

```{r paged.print=FALSE}
multi.model2 <- lm(Chance.of.Admit ~ CGPA + LOR + Research, data=dat)
multi.model2
summary(multi.model2)
supernova(multi.model2)
confint(multi.model2)
```
Extra Credit
```{r paged.print=FALSE}
q1 <- makeFun(multi.model)
q1
pred234 <- q1(dat[234,"CGPA"], dat[234, "LOR"], dat[234, "SOP"], dat[234, "Research"])
pred234

q2 <- resid(multi.model)
q2[128]
```