---
Author: "Ryan Paw"
title: "Assignment 2"
output: html_notebook
---
##SET UP
set working directory and load dataset
```{r}
myDataLocation <- "/Users/RyanPaw/My Files/R/ANA600/wk_2"

setwd(myDataLocation)

IBM_data <- read.csv(file = "IBM_data.csv", header = TRUE)
```
##SET UP
Next load favorite packages
```{r}
library(dplyr) 
library(ggplot2)
library(ggformula)
library(supernova)
library(lsr)
library(mosaic)
```
##NOTES - I do not want a word document. Use R Markdown and perform your work and writeup here. Maintain APA stype formatting for the parts where you are actually writing something down. 

##ACTIVITY
**Part 1. Purpose**
Describe three interesting hypotheses that you could ask with this data for one outcome (attrition, performance, age, etc.) variable. 

Each hypothesis should have one explanatory variable. **Your explanatory variables must include one quantitative variable and one categorical variable.** Write the relationships in word equations and give your interpretation for each.

1) What is the hourly rate (quant) of between different Genders (cat)? 
#Hourly Rate = Gender + Other Stuff
#My hypothesis is that Male employees might have a higher average hourly rate compared to Female employees. 
2) What is the hourly rate (quant) for employees by job role (cat)?
#Hourly Rate = Job Role + Other Stuff
#My hypothesis is the hourly rate for employees may increase for higher positioned job roles.
3) What is the hourly rate (quant) for employees with different job satisfaction(cat, num)?
#Hourly Rate = Job Satisfaction + Other Stuff
#My hypothesis is employees with higher pay may lead to higher job satisfaction.

**Part2: Data**
a) Describe the characteristics (demographics) of your sample (see Week One Assignment). 
    -This is where you are reviewing each variable and understanding whether it is cat or quant, and made up of numeric or str characters.
```{r}
str(IBM_data)

#HourlyRate = quantitative
#Gender = categorical
#Job Role = categorical
#Job Satisfaction = categorical, numeric 

```
b) Create a categorized variable from a quantitative variable that describes something about your sample, and include it in your description. (perhaps age or salary or categories) 
```{r}
#I categorized Job Satisfaction by converting 1,2,3,4 to Low, Medium, High, Very High. This new column is called "JobSatisfactionTier"
IBM_data$JobSatisfactionTier <- factor(IBM_data$JobSatisfaction,levels=c(1:4),labels=c("Low","Medium","High","Very High"))
```
c) Describe the criteria you used to define the categories. 
```{r}
#The criteria I used to define JobSatisfaction was to define each of the available numbers (1,2,3,4) with a rating scale. This will help create a categorical variable to better visualize JobSatisfaction. 
```
d) Report descriptive statistics **str(myData$variable)** and five-number summaries **fav_stats(myData$quantvar)** of your (1) outcome and (3) explanatory variables.
    - remember favstats is only for numerics (quant vars), and tally() or boxplots can be used for categorical vars.
```{r}
str(IBM_data$HourlyRate)
str(IBM_data$JobRole)
str(IBM_data$JobSatisfaction)
str(IBM_data$Gender)

favstats(IBM_data$HourlyRate)
summary(IBM_data$JobRole)
tally(IBM_data$JobSatisfaction)
summary(IBM_data$Gender)
```
e) Remove outliers and describe how and why you removed them.
```{r}
#To find outliers that are too low, we use the following equation: Q1 - 1.5*(Q3-Q1)
#To find outliers that are too high, we use the following equation: Q3 + 1.5*(Q3-Q1)
#To find the < outliers for HourlyRate:
LowOutliers <- 48-1.5*(83.75-48)
#To find the > outliers for HourlyRate:
HighOutliers <- 83.75+1.5*(83.75-48)
#I removed outliers in the HourlyRate column by using the subset() function. I named the new dataframe "NoOutliers". Based on my calculations for low and high outliers for Hourly Rate, no outliers were found. 
NoOutliers <- subset(IBM_data, (IBM_data$HourlyRate > LowOutliers) & (IBM_data$HourlyRate < HighOutliers))
NoOutliers
```
**Part3: Visualizations**
Create histograms, boxplots, jitter plots, or scatter plots, depending on whether they are quantitative/categorical and explanatory/outcome, that would best visualize your three hypotheses (one plot each). Use proper conventions and labels for your axis’s. Use smooth density overlays and facet_grids when applicable. **Be sure to use titles, axis labels, and legend labels if applicable.**
```{r}
gf_dhistogram(~HourlyRate,data=IBM_data,
             fill="darkgreen",color="black", 
             xlab="Employee Hourly Rate",ylab="Density of Employees",
             title="Employee Hourly Rate Between Different Genders") %>%
  gf_facet_grid(.~Gender) %>%
  gf_density()

gf_boxplot(HourlyRate~JobRole,data=IBM_data,
           xlab="Job Roles", ylab="Employee Hourly Rate",
           title="Employee Hourly Rate Between Job Roles")

gf_boxplot(HourlyRate~JobSatisfactionTier,data=IBM_data,
           xlab="Job Satisfaction",ylab="Employee Hourly Rate",
           title="Employee Hourly Rate Between Job Satisfaction")
```
**Part4: Results**
Based on your plots and analysis of the data, 
  a. Describe what the bars/points, x-axis, and y-axis mean in your plots. Be sure to give context by         referencing your variables (what they mean!). Reference the figures you created in Q3 as you discuss      them.
```{r}
#In the "Employee Hourly Rate Between Different Genders" histogram, the x-axis shows the Employee Hourly Rate and the y-axis shows the Density of Employees for each Hourly Rate. Each histogram is divided by Gender (Male/Female). 

#In the "Employee Hourly Rate Between Job Roles" boxplot, the x-axis shows the different Job Roles in the dataset (9 Job Roles total) and the y-axis shows the Employee Hourly Rate. This boxplot helps depict the different averages between each job role at the company. 

#In the "Employee Hourly Rate Between Job Satisfaction" boxplot, the x-axis shows the 4 different tiers of Job Satisfaction and y-axis shows the Employee Hourly Rate. This boxplot helps depict the average hourly rate with employee job satisfaction 
```
  b. What can you say from the distributions of your variables about their different characteristics     
     (e.g., shape, center, spread, kurtosis, skew, and outliers)?
```{r}
#In the "Employee Hourly Rate Between Different Genders" histogram, the shape of the Female plot is roughly uniform between each hourly rate. The shape of the Male plot has slightly more variation compared to the Female plot, but it is more or less uniform except a small valley around the $70 hourly rate. On both Male and Female plots, the distribution is about even. 

#In the "Employee Hourly Rate Between Job Roles" boxplot, the averages between each job role are roughly similar. The average between Human Resources and the other job roles is less. This shows that the Human Resources distribution is skewed right. All the remaining boxplots have a similar median line in the middle of the box. No outliers depicted for any of the boxplots. 

#In the "Employee Hourly Rate Between Job Satisfaction" boxplot, the median line between each box are similar. The height of the each boxplot are about the same. No skew for any boxplots (slight left skew for Low job satisfaction). No outliers depicted for any of the boxplots. 
```  
  c. What are your conclusions about the relationships between the outcome variable with the explanatory       variables for your three hypotheses?
```{r}
#"Employee Hourly Rate Between Different Genders": Male and females have about the same employee hourly rate distribution between the two. 

#"Employee Hourly Rate Between Job Roles": All job roles at IBM, except Human Resources (skewed right), have an evenly distributed hourly rate. 

#"Employee Hourly Rate Between Job Satisfaction": Job satisfaction is not significantly impacted by an employee's hourly rate. 
```
  d. What can you say about the population sampled, and the data-generating process?
```{r}
#Based on my conclusions, the population I sampled in this IBM dataset did not have as much variance as I expected. There may be other factors in the data-generating process that may be better suited to compare. The data generating process is the rules with which the data has been generated, and this dataset cleanly represents employee data at IBM based on our bottom-up approach.

```
Generalize your inferences from this sample to the intended population using APA style reporting.
#From my sample, I believe that there are other factors explaining the hourly rate at IBM. My results comparing Hourly Rate to Gender, Job Roles, and Job Satisfaction did not show as much variance as I expected, which makes me think about using other explanatory variables, such as Education and or PercentSalaryHike, to see if they have more significance.








**Step 5: Implications**
For one of your hypotheses, give interesting examples that could result in measurement error, sampling error, and mistakes. What are other variables in the dataset that could help explain variation in your outcome? What other variables, not in the dataset, could also explain variation in the outcome? How so? Are there any confounding variables that may change your interpretations?
#A sampling error that may have occurred is if person compiling this dataset may have not have understood which employees should be included/not included. Therefore, some employees may be not sampled, which could affect the results of this analysis. 

#A measurement error that could have occurred is if the "DistanceFromHome" variable was calcuated different. For example, Google Maps may say an employee is 10 miles away from the job to his/her house, but another software might say the same employee is 11 miles away from his/her house. This could affect any analyses that includes the "DistanceFromHome" variable. 

#Variation induced from mistakes could have affected the hourly rate. For example, if the person inputting an employee's hourly rate in the computer inputted the wrong number, it could affect the results of this sample. If an employee did not answer truthfully for any of the variables, it could significantly affect the results too. 

#I think adding a score for "Hazardous Working Conditions" could also explain variation in an outcome for hourly pay. Dangerous working conditions can be defined to include anything from handling dangerous chemicals in a research facility to walking in a dangerous job location. If there was a score, maybe from "1-10", that showed if there are any hazardous working conditions it could affect the outcome. Hazardous conditions could also be a confounding variable that could affect, for example, hourly pay and gender. A more dangerous workplace could create higher pay and more males may take the job than females.



 **Step 6: Discussion**
To recap, how could your analysis report be used by an employer? What do you believe is the practical significance of your analysis? Why is it important to understand the nature and circumstances that could influence the relationship between your outcome and explanatory variables?
#My analysis could be used by an employer if they wanted to know if hourly pay is affected between gender, job roles, and job satisfaction. However, the significance of my results are minimal, and I do not think my results show high significance between my chosen variables. I should have chosen different explanatory variables, such as Education or PercentSalary Hike, and maybe I could have more practical significance between hourly pay. It's important to understand the nature and circumstances because an analysis report can affect the people behind the data. For example, this dataset is about employees at IBM and if an analysis shows high signifiance between hourly pay with some kind of explanatory variable, IBM may want to cut costs and potentially lay people off. 






