---
title: "Week 1, Assignment1"
author: "Ryan Paw"
date: "10/3/2020"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#LOAD DATA
myDataLocation <- "/Users/RyanPaw/My Files/R/ANA605/wk_1"
setwd(myDataLocation) 
MyData<- read.csv(file = file.path(myDataLocation, "admission_data.csv"), header = TRUE)

#LOAD PACKAGES
library(dplyr) 
library(ggplot2)
library(ggformula)
library(supernova)
library(lsr)
library(mosaic)
library(skimr)
library(stargazer)
library(robust)
library(corrplot)

```
```{r}
str(MyData) 
```
Q1: 1.	For each of the following explanatory variables, fit a separate regression and correlation for the outcome variable, "chance of admit"


```{r}
#Factor the Binary numerics and plot using boxplot
MyData$ResearchFactor <- factor(MyData$Research, levels=c(0,1),labels=c("No","Yes"))
gf_boxplot(Chance.of.Admit~ResearchFactor,data=MyData)

#Regression
research.model <- lm(Chance.of.Admit~ResearchFactor,data=MyData)
research.model

#Correlation
research.correlation <- cor(Chance.of.Admit~Research,data=MyData)
research.correlation

supernova(research.model)

```
GPA
```{r}
#Plot
gf_point(Chance.of.Admit~CGPA,data=MyData) %>%
  gf_lm()

#Regression
GPA.model <- lm(Chance.of.Admit~CGPA,data=MyData)
GPA.model

#Correlation
GPA.correlation <- cor(Chance.of.Admit~CGPA,data=MyData)
GPA.correlation

supernova(GPA.model)

```
GRE
```{r}
#Plot
gf_point(Chance.of.Admit~GRE.Score,data=MyData) %>%
   gf_lm()
#Regression
GRE.model <- lm(Chance.of.Admit~GRE.Score,data=MyData)
GRE.model

#Correlation
GRE.correlation <- cor(Chance.of.Admit~GRE.Score,data=MyData)
GRE.correlation

supernova(GRE.model)

```
TOEFL
```{r}
#Plot
gf_point(Chance.of.Admit~TOEFL.Score,data=MyData) %>%
   gf_lm()
#Regression
TOEFL.model <- lm(Chance.of.Admit~TOEFL.Score,data=MyData)
TOEFL.model

#Correlation
TOEFL.correlation <- cor(Chance.of.Admit~TOEFL.Score,data=MyData)
TOEFL.correlation

supernova(TOEFL.model)

```
Q6: Obtain predictions for the following respondents for each of the explanatory variables below. Based on your analyses, who do you believe has the best chance of getting admitted? Why?

```{r}
#Make functions of your regression models
research.fun <- makeFun(research.model)
GPA.fun <- makeFun(GPA.model)
GRE.fun <- makeFun(GRE.model)
TOEFL.fun <- makeFun(TOEFL.model)

MyData$research.pred <- predict(research.model)
MyData$GPA.pred <- predict(GPA.model)
MyData$GRE.pred <- predict(GRE.model)
MyData$TOEFL.pred <- predict(TOEFL.model)

#Out of the 4 Respondents, I believe Respondent #56 has the best chance of getting admitted because his/her GRE Scores are much higher than the others which gives us a predicted Chance of Admit at 76%. Respondent #78 comes close to #56 because #78 has the highest GPA (65%) and our previous analysis showed that GPA is the strongest explanatory variable. However, #56 has an exceptionally high GRE score and good scores for our other variables, which makes me believe #56 has a higher chance of getting admitted.  
```
Respondent40
```{r}
# Respondent #40
res40 <- MyData[40,]
research.fun(res40$Research)
GPA.fun(res40$CGPA)
GRE.fun(res40$GRE.Score)
TOEFL.fun(res40$TOEFL.Score)


```
Other Respondents
```{r}
# Other respondents (substitute in with dat40)
#Respondent #56
res56 <- MyData[56,]
research.fun(res56$Research)
GPA.fun(res56$CGPA)
GRE.fun(res56$GRE.Score)
TOEFL.fun(res56$TOEFL.Score)


#Respondent #78
res78 <- MyData[78,]
research.fun(res78$Research)
GPA.fun(res78$CGPA)
GRE.fun(res78$GRE.Score)
TOEFL.fun(res78$TOEFL.Score)


#Respondent #93
res93 <- MyData[93,]
research.fun(res93$Research)
GPA.fun(res93$CGPA)
GRE.fun(res93$GRE.Score)
TOEFL.fun(res93$TOEFL.Score)


```
Extra Credit
```{r}
#GPA Model
#Yi=-1.07+0.21Xi+ei    
#i=100

MyData$GPA.resid <- resid(GPA.model)
MyData[100,"GPA.resid"]

Yi = 0.79
b0 = -1.07
b1 = 0.21
Xi = 8.88 
ei = 6.948265e-03 = 0.00

-1.07 + (0.21*8.88) + 0.01

```